---
title: "Football Transfer Window Analysis"
author: "Nassos Stylianou"
date: "October 2017"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
require(RCurl)
require(dplyr)
require(ggplot2)

```

First up, we need to get the transfer data, which [Tom Worville](https://twitter.com/Worville) has very kindly put online on his [Github page](https://github.com/Worville/guardian_transfers/blob/master/guardian_all_transfers_2017-09-02.csv)

```{r}
transfer_data <- read.csv(text=getURL("https://raw.githubusercontent.com/Worville/guardian_transfers/master/guardian_all_transfers_2017-09-02.csv"), header=T, stringsAsFactors = FALSE)

```
This is how the first few rows of the data look like: 

```{r}
head(transfer_data)
```
The first step is to make a new dataframe groupping the data by new club, ordering it by the most incoming transfers per club. 

```{r} 
incoming_transfers <- group_by(transfer_data, new_club) %>% summarise(count = n()) %>% arrange(desc(count))
names(incoming_transfers) <- c("new_club", "players_in")
head(incoming_transfers, 10)
```
Next step is to make a new dataframe groupping the data by the player's previous club, ordered by the club with most outgoing transfers

```{r}
outgoing_transfers <- group_by(transfer_data, previous_club) %>% summarise(count = n()) %>% arrange(desc(count))
names(outgoing_transfers) <- c("club", "players_out")
top_10 <- head(outgoing_transfers, 10)
```

Groups by the cost of incoming transfers by club and orders dataframe by club with most spent to buy players
```{r}
incoming_transfers_cost <- group_by(transfer_data, new_club) %>% summarise(total = sum(price_pounds, na.rm=TRUE)) %>% arrange(desc(total))
head(incoming_transfers_cost, 10)
ggplot
```
Groups by the cost of outgoing transfers by club, ordered by  club with most money received selling players 

```{r}
outgoing_transfers_cost <- group_by(transfer_data, previous_club) %>% summarise(total = sum(price_pounds, na.rm=TRUE)) %>% arrange(desc(total))
head(outgoing_transfers_cost, 10)
```

#merges cost of incoming transfers and outgoing transfer dataset and creates a column on total net spend by club
net_spend_per_club <- merge(incoming_transfers_cost, outgoing_transfers_cost, by.x=c("new_club"), by.y=c("previous_club")) %>% mutate(net_spend = total.x - total.y) %>% select(new_club, net_spend)

#creates a unique list of all clubs in the data 
new_clubs <- select(transfer_data, new_club, new_league) 
names(new_clubs) <- c("club", "league")

old_clubs <-select(transfer_data, previous_club, previous_league)
names(old_clubs) <- c("club", "league")

#removes clubs which appear twice
all_clubs <- rbind(new_clubs, old_clubs) %>% subset(!duplicated(club))

#merges club list with net spend dataset and keeps clubs only in the big five European leagues. Removes Pescara which looks like an error and adds a column each for total incoming and total outgoing players
net_spend_per_club_top_five_leagues <- merge(net_spend_per_club, all_clubs, by.x = "new_club", by.y = "club") %>% filter(league == "Premier League" | league == "Serie A" | league == "Bundesliga" | league == "La Liga" | league == "Ligue 1") %>% arrange(desc(league)) %>% filter(new_club != "Pescara") %>% merge(incoming_transfers, by="new_club") %>% merge(outgoing_transfers, by.x = "new_club", by.y = "club") %>% View() %>% write.csv("data.csv")

group_by(net_spend_per_club_top_five_leagues, league) %>% summarise(count = n())

